Informing forest carbon inventories under the Paris Agreement using the Global Forest Carbon Database (ForC v4.0)

### 2023-05-23 doc/manuscript/ESSDmanuscript.tex
incorporate Sandro's comments (send in Word doc)
plus  a few other changes, including title

#39

\title{Informing IPCC accounting of forest carbon using the global
forest carbon database (ForC v4.0)}

@@

\textbf{Abstract.} Forests are critical for climate change mitigation
and constitute a substantial portion of planned emissions reductions
under the 2015 Paris Agreement. Yet, the efficacy of greenhouse gas
mitigation planning and reporting is dependent upon the quality of
available emission factors data, including forest carbon (C) stocks and
changes therein. Tens of thousands of relevant forest C estimates have
been published, yet are not readily accesible to the practitioners
compiling national greenhouse gas inventories. Many of these data have,
however, been compiled in the Global Forest C database (ForC;
https://forc-db.github.io/) and stand to be of value to greenhouse gas
accounting if made available through the Emission Factor Database (EFDB)
of the International Panel on Climate Change (IPCC). Here, we develop
and document a process for semi-automated transfer of data from ForC
into the EFDB, assess the data available and transferred to date, and
provide recommendations for improving forest data collection, analysis,
and reporting to improve accounting of forest-sector greenhoouse gas
emissions and removals. We begin by reconciling terminology and mapping
ForC fields into EFDB. This process required some updates to the ForC
database structure, leading to the release of a new version of ForC
(v4.0; described here). As of May 16, 2023, ForC contained
\textasciitilde19323 independent records relevant to EFDB, 1438 of which
have been submitted to date. Among the data in ForC, there is
disproportionate representation of biomass (particularly aboveground)
stocks, with far fewer records for dead organic matter and soil C, and
relatively few or no records for net annual increments or C fluxes into
(gains) or out of (losses) the IPCC-defined C pools. Geographic
representation is also quite uneven, with the highest densities of
relevant records in temperate forests, and with relatively scant
representation of tropical forests in Africa and Asia. This distribution
of records is generally reflected in the subset of records that have
been submitted to EFDB to date. In the future, forest C estimates in
EFDB can be improved through targeted research to fill critical gaps,
reporting of information required by IPCC, and continued submission of
data from scientific publications to the EFDB. Given that climate change
is rapidly impacting the world's forests, timely reporting of recent
estimates will be especially critical to accurate forest C accounting.

@@

Forests are critical to management of atmospheric concentrations of the
greenhouse gas carbon dioxide (CO\textsubscript{2}), and thereby climate
change. In recent decades, CO\textsubscript{2} uptake by forests,
woodlands, and savannas has exceeded releases from deforestation and
other severe disturbances, resulting in a net carbon CO\textsubscript{2}
sink of \textasciitilde0.88 Gt C yr\textsuperscript{-1} \citep[all
biomes with trees,][]{xu_changes_2021} to \textasciitilde1.6 Gt C
yr\textsuperscript{-1} \citep[forests only,][]{harris_global_2021}. This
has offset an estimated 10\% to 18\% of anthropogenic
CO\textsubscript{2} emissions from fossil fuels and cement
\citep{xu_changes_2021, harris_global_2021}, dramatically slowing the
pace of atmospheric CO\textsubscript{2} accumulation and climate change.
Going into the future, the fate of this important CO\textsubscript{2}
sink is highly uncertain, depending both upon forest responses to
climate change, which are likely to reduce the sink strength
\citep{mcdowell_pervasive_2020, hammond_global_2022}, and on human
conservation, restoration, and management of forests
\citep{ipcc_climate_2019, ipcc_climate_2022}.

@@

The International Panel on Climate Change (IPCC) provides guidance for
national greenhouse gas inventories for reporting to the United Nations
Framework Convention on Climate Change
\citep[UNFCCC,][]{ipcc_2006_2006, ipcc_2019_2019}. Under this guidance,
greenhouse gas inventories include all managed land, including most of
the world's forest land \citep{ogle_delineating_2018}. The IPCC
inventory guidelines include specific instructions for accounting for
greenhouse gas (mainly CO\textsubscript{2}) exchanges between forest
land and the atmosphere \citep{ipcc_agriculture_2006, ipcc_2019_2019}. A
tiered approach to accounting is employed, where the lowest tier (Tier
1) represents the simplest approach and relies on default parameter
values -- for example, forest carbon (C) stocks values by ecozone
\citep{fao_global_2012} and forest age class derived as the average of
published estimates \citep{ipcc_2019_2019, rozendaal_aboveground_2022}.
Tier 1 values have improved over the years as more of the relevant
underlying data has become available
\citep{requenasuarez_estimating_2019, rozendaal_aboveground_2022}, but
there remains room for continuous improvement as the science advances.
For example, the year following the release of the latest IPCC
guidelines, a more thorough analysis of C accumulation in regrowth
forests found that IPCC's Tier 1 default values underestimated C
sequestration by 32\% on average and failed to capture eight-fold
variation within ecozones \citep{cook-patton_mapping_2020}. In addition,
it was revealed that C stocks in mature African tropical montane forests
were two-thirds higher than the IPCC Tier 1 values for these forests
\citep{cuni-sanchez_high_2021}. This rapid evolution of scientific
information on C cycling in forests is valuable for informing climate
change mitigation efforts but requires improved mechanisms for
communicating the latest information from scientific researchers to the
practitioners who need reliable estimates for greenhouse gas mitigation
planning. Moreover, high variability of forest C cycling within ecozones
\citep[e.g.,][]{cook-patton_mapping_2020, cuni-sanchez_high_2021}
implies that it is useful for practitioners to have access to
locally-specific information, when available.

To improve data accessibility for C accounting, the IPCC created the
Emission Factor Database (EFDB;
\url{https://www.ipcc-nggip.iges.or.jp/EFDB/main.php}), which is
intended as a recognized library of emission factors and other
parameters that can be used for estimating greenhouse gas emissions and
removals. The EFDB can be used both for efforts to tally a nation's
intended or accomplished greenhouse gas reductions, or as a basis of
comparison for external parties to evaluate these inventories. The EFDB
encourages researchers to submit estimates of emission factors or other
related parameters from peer-reviewed journal papers or other accepted
sources for inclusion in the database. In the case of forests, emission
factors include C stocks, net increments (``stock changes''), and fluxes
(``gains'' and ``losses'') for various pools
\citep{ipcc_2006_2006, ipcc_2019_2019}.

The Global Forest Carbon Database, ForC
(\url{https://forc-db.github.io/}), is the largest collection of
published estimates of forest C stocks, increments, and annual fluxes
\citep{anderson-teixeira_forc_2018, anderson-teixeira_carbon_2021}. ForC
includes data ingested from individual publications and relevant
databases, including the Global Reforestation Opportunity Assessment
(GROA) database \citep[database doi:
10.5281/zenodo.3983644]{cook-patton_mapping_2020}, the global soil
respiration database
\citep[SRDB-V5,][]{bond-lamberty_global_2010, jian_restructured_2021}.
As of May 16, 2023, ForC contained 39848 records from 10589 plots in
1535 distinct geographical areas, along with records of stand age and
disturbance history. As such, ForC is positioned to improve forest C
accounting through the transfer of data to EFDB. The purpose of this
publication is to document that process and provide recommendations for
future improvements.

Here, we (1) review IPCC methods and definitions for forest C accounting
in the context of typical forest C estimation methodologies; (2)
describe mapping of ForC to IPCC's EFDB; (3) describe updates to ForC
(ForC v4.0), most of which were implemented to facilitate data transfer
to EFDB; (4) summarize the data in ForC that's relevant to EFDB and
records that have been transferred to date; and (5) provide
recommendations for improving data collection, analysis, database, and
accounting.

@@

Litter, which is typically \textless40 t C ha\textsuperscript{-1} but
can exceed 100 t C ha\textsuperscript{-1}
\citep{anderson-teixeira_carbon_2021}, is defined by IPCCC as including
``all non-living biomass with a diameter less than a minimum diameter
chosen by the country (for example 10 cm), lying dead, in various states
of decomposition above the mineral or organic soil''
\citep{ipcc_good_2003, ipcc_2006_2006}. As noted above, live fine roots
may be included in litter when difficult to separate empirically. The
definition includes the entire O horizion, including litter (OL), fumic
(OF), and humic (OH) layers, in addition to litter embedded within the
soil. This definition contrasts with empirical studies that focus on
aboveground litter, often including only the OL layer in the definition
of litter, and do not always specify the components included. Similar to
dead wood, litter is poorly characterized in large-scale forest C
budgets \citep{pan_large_2011, harris_global_2021}, and IPCC's latest
Tier 1 default values are based on just 1-7 references per climate zone
\citep[Table 2.2 in][]{ipcc_2019_2019}.

@@

Previous versions of ForC
\citep{anderson-teixeira_carbon_2016, anderson-teixeira_forc_2018, anderson-teixeira_carbon_2021}
contained most of the information required by EFDB, and, more broadly,
to inform C accounting under IPCC guidelines. However, modest changes to
the structure and contents of ForC were needed in order to provide all
information required by EFDB and to improve ForC's capacity to serve as
a repository of valuable information for forest C accounting under IPCC
guidelines. To support export of data to EFDB, and to improve the
overall quality of the ForC database, we added or modified 18 fields
(Appendix A), defined 15 new variables, implemented enhanced quality
control, manually reviewed \textgreater1954 records to obtain additional
required information, and added 329 new records.

@@

To create structure for EFDB-relevant records, we added a total of 15
new variables to the set of named and defined variables (Fig. 1),
counting each pair of variables with units in C (ending in \texttt{\_C})
or organic matter (ending in \texttt{\_OM}) as one. The majority of
these were increment variables (n=11), adding to only one previously
defined increment variable (aboveground biomass increment,
\emph{delta.agb}). These are directly related to C stocks as previously
defined in ForC, with ``\emph{delta.}'' added in front of the variable
name. Further, we added variables capturing the belowground component of
woody mortality (\emph{woody.mortality\_root}) and the combined
aboveground and belowground components of woody mortality
(\emph{woody.mortality}). Although most of these variables lacked
records in ForC as of May 16, 2023, their addition gave the structure
such that records can be populated over time. Finally, to provide better
definition of the previously existing variable \emph{organic.layer},
which has a nebulous definition that reflects the varied definitions
adopted by original studies, we added two clearly defined variables:
\emph{litter} (relatively undecomposed plant material/ OL horizon), and
\emph{O.horizon} (entire O-horizon, including OL).

@@

\section{Transfer of data from ForC to EFDB}

To transfer complete, reviewed ForC records into EFDB, we created R
scripts to restructure ForC records and populate EFDB's bulk import form
(``EFDB bulk import.xlsx''). Criteria for data transfer were that (1)
records had been checked against the original study and determined to be
complete and correct, and as originally presented, (2) the original
study presented values in tables or text, as opposed to the values
having been digitized from graphs or calculated based on related
variables, and (3) the records had not previously been sent to EFDB.
Once converted into EFDB format, the records were reviewed and then sent
to the IPCC's Technical Support Unit for inclusion in EFDB. Complete
records needed to be reviewed by the EFDB editorial board prior to
posting in the database -- a process that lags behind records transfer
and had not yet been completed for all records sent as of May 16, 2023.

@@

With input from the IPCC's Technical Support Unit, we reviewed the list
of ForC variables to identify those that were relevant to EFDB and to
appropriately map them into EFDB (Fig. 1). For each C pool (Table 1), we
identified variables representing organic matter or C stocks, stock
changes (a.k.a. ``net annual increments'' by IPCC, ``increments'' in
ForC), gains (a.k.a. ``gross annual increments'' by IPCC, ``fluxes'' in
ForC), and losses (``fluxes'' in ForC). As described in section 3.2, we
also defined 15 new EFDB-relevant variables that were not previously
represented in ForC. It it important to note that the correspondence of
ForC variables to IPCC criteria often depends upon measurement protocols
(``important sources of estimate variation'' in Table 1). For example,
ForC records of biomass and dead wood vary in the minimum stem diameter
censused, such that some records would match the IPCC criteria whereas
others would not. Information on minimum diameters censused and other
important sources of methodological variation are recorded as covariates
in ForC and mapped into the EFDB field \emph{Other Properties} (Appendix
B). Details on the mapping of ForC variables to EFDB -- including
associated covariates, IPCC pools (Table 1) and relevant equations
\citep{ipcc_2006_2006} -- are documented in the file
ForC\_variables\_mapping.csv in the GitHub repository associated with
this publication IPCC-EFDB-integration repository in ForC-db
organization (\url{https://github.com/forc-db/IPCC-EFDB-integration}).

@@

Classification into sub-categories was dependent upon stand age and site
history (section 2.2). For Forest Land ≥ 20 years old or of unknown
(relatively mature) age, or Forest Land \textless{} 20 years old that
was forest prior to a stand-clearing disturbance, the past land-use
category was Forest Land, making the sub-category ``Forest Land
Remaining Forest land''. For forests \textless20 years old with history
including cultivation/ tillage or grazing, past land-use categories were
Cropland and Grassland, respectively, making land-use subcategories were
``Cropland converted to Forest Land'' and ``Cropland converted to Forest
Land'', respectively. For forests \textless20 years old with unspecified
previous agricultural use, we assigned the sub-category ``Land Converted
to Forest land''. Forests \textless20 years old with unknown land use
prior to the study date were simply classified as ``Forest Land''. The
same logic was applied for savannas, but including both Forest Land and
Grassland as potentially relevant categories.

Given the lack of public information needed to determine whether lands
are classified as mangaged
\citep{ogle_delineating_2018, deng_comparing_2021}, and because the
IPCC's definition of managed land is more expansive than is commonly
applied in the scientific literature and hence in ForC, we did not
transfer any classification of land management status from ForC to the
EFDB. However, we do provide auxiliary information that should be useful
in making this determination, including geographical location and
notable disturbance events.

@@

As of May 16, 2023, ForC (v4.0) contained 32693 independent records
(39848 total), 19323 of which were for the the 43 variables relevant to
EFDB (Fig. 1). These records were distributed across all forested
continents and ecozones, albeit unevenly (Fig. 2). The largest number of
records came from Asia, followed by North America, South America, and
Europe, with relatively few records from Africa, Australia, and Oceania
(Fig. 3c). Categorized by FAO ecozone, the greatest numbers of records
came from subtropical humid forests, temperate mountain systems, and
tropical rain forests, each with \textgreater2,000 independent records
(Fig. 3b). Boreal coniferous forests, temperate continental forests,
subtropical mountain systems, and tropical moist deciduous forests had
\textgreater1,000 independent records each, while other ecozones all had
\textless1,000 records. The most widely represented forest type was
needleleaf evergreen, followed by broadleaf deciduous and broadleaf
evergreen (Fig. 3a). In terms of stand age, the most represented age
class was 20-100 years, followed by \textless20 years and then
\textgreater100 years (Fig. 3d).

@@

\subsection{Data transfers to EFDB}

As of May 16, 2023, we had reviewed or added 2283 EFDB-relevant records,
1438 records of which were sent to EFDB, and 376 of which have been
reviewed, accepted, and posted (Figs. 2-3, Table 2). The 37\%
attenuation between records reviewed and those sent to EFDB was
attributable to the presence of digitized records and records where a
variable's value had been calculated as the sum or difference of related
variables rather than presented directly in the text. The discrepancy
between the number of records sent and that posted to EFDB is primarily
attributable to the time required for the IPCC to review the records,
and it is not expected that many -- if any -- records will be rejected.

@@

There is vast potential to expand forest C data in EFDB by completing
the process of reviewing and sending data that are already in ForC
(Figs. 2-3). So far, only \textasciitilde7\% of the EFDB-relevant data
in ForC have been sent to EFDB. Although this process requires manual
review of records, the addition of new records to EFDB is hugely
facilitated by the fact that most pertinent information for each record
is already entered in ForC and can be easily transferred to EFDB using
the system developed here. Future efforts to review studies for transfer
should optimize for representation across geographic regions, forest
types, and variables, giving priority to those from currently
under-represented regions and forest types (Figs. 2-3, Table 2), to
records from countries relying on existing data for their greenhouse gas
inventories (Tier 1 or 2 methodology), to the variables most needed by
EFDB users, and to the more contemporary records.

@@

Filling knowledge gaps in other EFDB-relevant variables will require
more effort, but this effort is warranted given their importance for
forest C accounting. Although aboveground biomass is the most studied
variable considered here (Table 2) and is the target of satellite
missions
\citep{dubayah_global_2020, quegan_european_2019, nisar_nasaisro_2018},
significant ground-based research effort is required to create accurate
global maps of forest biomass and changes therein
\citep{duncanson_importance_2019, labrière_forest_2023}. Given
observations of increasing tree mortality in some forested regions
\citep{mcdowell_pervasive_2020}, better characterization of forest dead
wood will be critical. Additionally, C stocks in forest organic horizons
and soils can be quite substantial and highly uncertain in many parts of
the world \citep{tifafi_large_2018}. Significant investment in
ground-based forest research will be critical to filling these gaps.

@@

We recommend that, in order to make research valuable to C accounting
under IPCC guidelines, researchers calculate and report results
according to IPCC standards (Table 3). It is particularly noteworthy
that simple decisions on the presentation of results will determine
whether the records meet the criteria for inclusion in EFDB. Some
examples are as follows: (1) presenting data only in a figure makes them
ineligible for inclusion in EFDB, whereas presentation in a table or
supplementary data file allows inclusion; (2) direct presentation of all
relevant variables allows inclusion, whereas presenting only components
of variables of interest (e.g., parsing litter into fine woody debris,
OL, OF, and OH layers) or requiring simple mathematical operations to
obtain a variable of interest (e.g., \emph{delta.agb} =
\emph{ANPP\_woody} - \emph{woody.mortality.agb}) disqualifies records
from inclusion; (3) matching IPCC-defined thresholds for defining C
pools (Table 1), which may vary by country, can make the data far more
relevant for IPCC accounting (e.g., using a 10 cm cutoff between dead
wood and litter, presenting soil C to a depth of 30 cm). It should also
be emphasized that reporting of 95\% confidence intervals (or other
metrics of error), when applicable, is highly desirable and makes the
data more relevant to IPCC.

>>>>

\title{Informing forest carbon inventories under the Paris Agreement
using the Global Forest Carbon Database (ForC v4.0)}

@@

\textbf{Abstract.} Forests are critical for climate change mitigation
and constitute a substantial portion of planned net emissions reductions
under the 2015 Paris Agreement. Yet, the efficacy of greenhouse gas
mitigation planning and reporting is dependent upon the quality of
available emission factors data, including forest carbon (C) stocks and
changes therein. Tens of thousands of relevant forest C estimates have
been published, yet are not readily accesible to the practitioners
compiling national greenhouse gas inventories. Many of these data have,
however, been compiled in the Global Forest C database (ForC;
https://forc-db.github.io/) and stand to be of value to greenhouse gas
accounting if made available through the Emission Factor Database (EFDB)
of the International Panel on Climate Change (IPCC). Here, we develop
and document a process for semi-automated submission of data from ForC
into the EFDB, assess the data available and transferred to date, and
provide recommendations for improving forest data collection, analysis,
and reporting to improve accounting of forest-sector greenhouse gas
emissions and removals. We begin by reconciling terminology and mapping
ForC fields into EFDB. This process required some updates to the ForC
database structure, leading to the release of a new version of ForC
(v4.0; described here). As of May 23, 2023, ForC contained
\textasciitilde19323 independent records relevant to EFDB, 1438 of which
have been submitted to date. Among the data in ForC, there is
disproportionate representation of biomass (particularly aboveground)
stocks, with far fewer records for dead organic matter and soil C, and
relatively few or no records for net annual increments or C fluxes into
(gains) or out of (losses) the IPCC-defined C pools. Geographic
representation is also quite uneven, with the highest densities of
relevant records in temperate forests, and with relatively scant
representation of tropical forests in Africa and Asia. This distribution
of records is generally reflected in the subset of records that have
been submitted to EFDB to date. In the future, forest C estimates in
EFDB can be improved through targeted research to fill critical gaps,
reporting of information required by IPCC, and continued submission of
data from scientific publications to the EFDB. Given that climate change
is rapidly impacting the world's forests, timely reporting of recent
estimates will be especially critical to accurate forest C accounting.

@@

Forests are critical to management of the atmospheric concentration of
the greenhouse gas carbon dioxide (CO\textsubscript{2}), and thereby
climate change. In recent decades, CO\textsubscript{2} uptake by
forests, woodlands, and savannas has exceeded releases from
deforestation and other severe disturbances, resulting in a net carbon
CO\textsubscript{2} sink of \textasciitilde0.88 Gt C
yr\textsuperscript{-1} \citep[all biomes with trees,][]{xu_changes_2021}
to \textasciitilde1.6 Gt C yr\textsuperscript{-1} \citep[forests
only,][]{harris_global_2021}. This has offset an estimated 10\% to 18\%
of anthropogenic CO\textsubscript{2} emissions from fossil fuels and
cement \citep{xu_changes_2021, harris_global_2021}, dramatically slowing
the pace of atmospheric CO\textsubscript{2} accumulation and associated
climate change. Going into the future, the fate of this important
CO\textsubscript{2} sink is highly uncertain, depending both upon forest
responses to climate change, which are likely to reduce the sink
strength \citep{mcdowell_pervasive_2020, hammond_global_2022}, and on
human conservation, restoration, and management of forests
\citep{ipcc_climate_2019, ipcc_climate_2022}.

@@

The International Panel on Climate Change (IPCC) provides guidance for
national greenhouse gas inventories for reporting to the United Nations
Framework Convention on Climate Change
\citep[UNFCCC,][]{ipcc_2006_2006, ipcc_2019_2019}. Under this guidance,
greenhouse gas inventories include all managed land, including most of
the world's forest land \citep{ogle_delineating_2018}. The IPCC
inventory guidelines include specific instructions for accounting for
greenhouse gas (mainly CO\textsubscript{2}) exchanges between forest
land and the atmosphere \citep{ipcc_agriculture_2006, ipcc_2019_2019}. A
tiered approach to accounting is employed, where the lowest tier (Tier
1) represents the simplest approach and relies on default parameter
values -- for example, forest carbon (C) stocks values by ecozone
\citep{fao_global_2012} and forest age class derived as the average of
published estimates \citep{ipcc_2019_2019, rozendaal_aboveground_2022}.
Tier 1 values have improved over the years as more of the relevant
underlying data has become available
\citep{requenasuarez_estimating_2019, rozendaal_aboveground_2022}, but
there remains room for continuous improvement as the science advances
and datasets grow and become more widely accessible. For example, the
year following the release of the latest IPCC guidelines, a more
thorough analysis of C accumulation in regrowth forests found that
IPCC's Tier 1 default values underestimated C sequestration by 32\% on
average and failed to capture eight-fold variation within ecozones
\citep{cook-patton_mapping_2020}. In addition, it was revealed that C
stocks in mature African tropical montane forests were two-thirds higher
than the IPCC Tier 1 values for these forests
\citep{cuni-sanchez_high_2021}. This rapid evolution of scientific
information on C cycling in forests is valuable for informing climate
change mitigation efforts but requires improved mechanisms for
communicating the latest information from scientific researchers to the
practitioners who need reliable estimates for greenhouse gas mitigation
planning. Moreover, high variability of forest C cycling within ecozones
\citep[e.g.,][]{cook-patton_mapping_2020, cuni-sanchez_high_2021}
implies that it is useful for practitioners to have access to
locally-specific information, when available.

To improve data accessibility for preparing greenhouse gas estimates,
the IPCC created the Emission Factor Database (EFDB;
\url{https://www.ipcc-nggip.iges.or.jp/EFDB/main.php}), which is
intended as a recognized library of emission factors and other
parameters that can be used for estimating greenhouse gas emissions and
removals. The EFDB can be used both for efforts to tally a nation's
intended or accomplished greenhouse gas reductions, or as a basis of
comparison for external parties to evaluate these inventories. The EFDB
encourages researchers to submit estimates of emission factors or other
related parameters from peer-reviewed journal papers or other accepted
sources for inclusion in the database. In the case of forests, emission
factors include C stocks, net increments (``stock changes''), and fluxes
(``gains'' and ``losses'') for various pools
\citep{ipcc_2006_2006, ipcc_2019_2019}.

The Global Forest Carbon Database, ForC
(\url{https://forc-db.github.io/}), is the largest collection of
published estimates of forest C stocks, increments, and annual fluxes
\citep{anderson-teixeira_forc_2018, anderson-teixeira_carbon_2021}. ForC
includes data ingested from individual publications and relevant
databases, including the Global Reforestation Opportunity Assessment
(GROA) database \citep[database doi:
10.5281/zenodo.3983644]{cook-patton_mapping_2020}, the global soil
respiration database
\citep[SRDB-V5,][]{bond-lamberty_global_2010, jian_restructured_2021}.
As of May 23, 2023, ForC contained 39848 records from 10589 plots in
1535 distinct geographical areas, along with records of stand age and
disturbance history. As such, ForC is positioned to improve
forest-related estimates of CO\textsubscript{2} emissions and removals
through the submission of data to the EFDB. The purpose of this
publication is to document that process and provide recommendations for
future improvements.

Here, we (1) review IPCC methods and definitions applied to estimate CO2
emissions and removals from forest in the context of typical forest C
estimation methodologies; (2) describe mapping of ForC to IPCC's EFDB;
(3) describe updates to ForC (ForC v4.0), most of which were implemented
to facilitate data transfer to EFDB; (4) summarize the data in ForC
that's relevant to EFDB and records that have been submitted to date;
and (5) provide recommendations for improving data collection, analysis,
database, and accounting.

@@

Litter, which is typically \textless40 t C ha\textsuperscript{-1} but
can exceed 100 t C ha\textsuperscript{-1}
\citep{anderson-teixeira_carbon_2021}, is defined by IPCC as including
``all non-living biomass with a diameter less than a minimum diameter
chosen by the country (for example 10 cm), lying dead, in various states
of decomposition above the mineral or organic soil''
\citep{ipcc_good_2003, ipcc_2006_2006}. As noted above, live fine roots
may be included in litter when difficult to separate empirically. The
definition includes the entire O horizion, including litter (OL), fumic
(OF), and humic (OH) layers, in addition to litter embedded within the
soil. This definition contrasts with empirical studies that focus on
aboveground litter, often including only the OL layer in the definition
of litter, and do not always specify the components included. Similar to
dead wood, litter is poorly characterized in large-scale forest C
budgets \citep{pan_large_2011, harris_global_2021}, and IPCC's latest
Tier 1 default values are based on just 1-7 references per climate zone
\citep[Table 2.2 in][]{ipcc_2019_2019}.

@@

Previous versions of ForC
\citep{anderson-teixeira_carbon_2016, anderson-teixeira_forc_2018, anderson-teixeira_carbon_2021}
contained most of the information required by EFDB, and, more broadly,
to inform C stock change calculations under the Paris Agreement.
However, modest changes to the structure and contents of ForC were
needed in order to provide all information required by EFDB and to
improve ForC's capacity to serve as a repository of valuable information
for forest C inventories under IPCC guidelines. To support export of
data to EFDB, and to improve the overall quality of the ForC database,
we added or modified 18 fields (Appendix A), defined 15 new variables,
implemented enhanced quality control, manually reviewed \textgreater1954
records to obtain additional required information, and added 329 new
records.

@@

To create structure for EFDB-relevant records, we added a total of 15
new variables to the set of named and defined variables (Fig. 1),
counting each pair of variables with units in C (ending in \texttt{\_C})
or organic matter (ending in \texttt{\_OM}) as one. The majority of
these were increment variables (n=11), adding to only one previously
defined increment variable (aboveground biomass increment,
\emph{delta.agb}). These are directly related to C stocks as previously
defined in ForC, with ``\emph{delta.}'' added in front of the variable
name. Further, we added variables capturing the belowground component of
woody mortality (\emph{woody.mortality\_root}) and the combined
aboveground and belowground components of woody mortality
(\emph{woody.mortality}). Although most of these variables lacked
records in ForC as of May 23, 2023, their addition gave the structure
such that records can be populated over time. Finally, to provide better
definition of the previously existing variable \emph{organic.layer},
which has a nebulous definition that reflects the varied definitions
adopted by original studies, we added two clearly defined variables:
\emph{litter} (relatively undecomposed plant material/ OL horizon), and
\emph{O.horizon} (entire O-horizon, including OL).

@@

\section{Submission of ForC data to EFDB}

To submit complete, reviewed ForC records into EFDB, we created R
scripts to restructure ForC records and populate EFDB's bulk import form
(``EFDB bulk import.xlsx''). Criteria for data submission were that (1)
records had been checked against the original study and determined to be
complete and correct, and as originally presented, (2) the original
study presented values in tables or text, as opposed to the values
having been digitized from graphs or calculated based on related
variables, and (3) the records had not previously been sent to EFDB.
Once converted into EFDB format, the records were reviewed and then sent
to the IPCC's Technical Support Unit for inclusion in EFDB. Complete
records needed to be reviewed by the EFDB editorial board prior to
posting in the database -- a process that lags behind records transfer
and had not yet been completed for all records sent as of May 23, 2023.

@@

With input from the IPCC's Technical Support Unit, we reviewed the list
of ForC variables to identify those that were relevant to EFDB and to
appropriately map them into EFDB (Fig. 1). For each C pool (Table 1), we
identified variables representing organic matter or C stocks, stock
changes (a.k.a. ``net annual increments'' by IPCC, ``increments'' in
ForC), gains (a.k.a. ``gross annual increments'' by IPCC, ``fluxes'' in
ForC), and losses (``fluxes'' in ForC). As described in section 3.2, we
also defined 15 new EFDB-relevant variables that were not previously
represented in ForC. It is important to note that the correspondence of
ForC variables to IPCC criteria often depends upon measurement protocols
(``important sources of estimate variation'' in Table 1). For example,
ForC records of biomass and dead wood vary in the minimum stem diameter
censused, such that some records would match the IPCC criteria whereas
others would not. Information on minimum diameters censused and other
important sources of methodological variation are recorded as covariates
in ForC and mapped into the EFDB field \emph{Other Properties} (Appendix
B). Details on the mapping of ForC variables to EFDB -- including
associated covariates, IPCC pools (Table 1) and relevant equations
\citep{ipcc_2006_2006} -- are documented in the file
ForC\_variables\_mapping.csv in the GitHub repository associated with
this publication IPCC-EFDB-integration repository in ForC-db
organization (\url{https://github.com/forc-db/IPCC-EFDB-integration}).

@@

Classification into sub-categories was dependent upon stand age and site
history (section 2.2). For Forest Land ≥ 20 years old or of unknown
(relatively mature) age, or Forest Land \textless{} 20 years old that
was forest prior to a stand-clearing disturbance, the past land-use
category was Forest Land, making the sub-category ``Forest Land
Remaining Forest land''. For forests \textless20 years old with history
including cultivation/ tillage or grazing, past land-use categories were
Cropland and Grassland, respectively, making land-use subcategories were
``Cropland converted to Forest Land'' and ``Grassland converted to
Forest Land'', respectively. For forests \textless20 years old with
unspecified previous agricultural use, we assigned the sub-category
``Land Converted to Forest land''. Forests \textless20 years old with
unknown land use prior to the study date were simply classified as
``Forest Land''. The same logic was applied for savannas, but including
both Forest Land and Grassland as potentially relevant categories.

Given the lack of public information needed to determine whether lands
are classified as managed
\citep{ogle_delineating_2018, deng_comparing_2021}, and because the
IPCC's definition of managed land is more expansive than is commonly
applied in the scientific literature and hence in ForC, we did not
transfer any classification of land management status from ForC to the
EFDB. However, we do provide auxiliary information that should be useful
in making this determination, including geographical location and
notable disturbance events.

@@

As of May 23, 2023, ForC (v4.0) contained 32693 independent records
(39848 total), 19323 of which were for the 43 variables relevant to EFDB
(Fig. 1). These records were distributed across all forested continents
and ecozones, albeit unevenly (Fig. 2). The largest number of records
came from Asia, followed by North America, South America, and Europe,
with relatively few records from Africa, Australia, and Oceania (Fig.
3c). Categorized by FAO ecozone, the greatest numbers of records came
from subtropical humid forests, temperate mountain systems, and tropical
rain forests, each with \textgreater2,000 independent records (Fig. 3b).
Boreal coniferous forests, temperate continental forests, subtropical
mountain systems, and tropical moist deciduous forests had
\textgreater1,000 independent records each, while other ecozones all had
\textless1,000 records. The most widely represented forest type was
needleleaf evergreen, followed by broadleaf deciduous and broadleaf
evergreen (Fig. 3a). In terms of stand age, the most represented age
class was 20-100 years, followed by \textless20 years and then
\textgreater100 years (Fig. 3d).

@@

\subsection{Data submissions to EFDB}

As of May 23, 2023, we had reviewed or added 2283 EFDB-relevant records,
1438 records of which were submitted to EFDB, and 376 of which have been
reviewed, accepted, and posted (Figs. 2-3, Table 2). The 37\%
attenuation between records reviewed and those sent to EFDB was
attributable to the presence of digitized records and records where a
variable's value had been calculated as the sum or difference of related
variables rather than presented directly in the text. The discrepancy
between the number of records sent and that posted to EFDB is primarily
attributable to the time required for the IPCC to review the records,
and it is not expected that many -- if any -- records will be rejected.

@@

There is vast potential to expand forest C data in EFDB by completing
the process of reviewing and submitting data that are already in ForC
(Figs. 2-3). So far, only \textasciitilde7\% of the EFDB-relevant data
in ForC have been submitted to EFDB. Although this process requires
manual review of records, the submission of new records to EFDB is
hugely facilitated by the fact that most pertinent information for each
record is already entered in ForC and can be easily prepared for
submission to EFDB using the system developed here. Future efforts to
review studies for transfer should optimize for representation across
geographic regions, forest types, and variables, giving priority to
those from currently under-represented regions and forest types (Figs.
2-3, Table 2), to records from countries relying on existing data for
their greenhouse gas inventories (Tier 1 or 2 methodology), to the
variables most needed by EFDB users, and to the more contemporary
records.

@@

Filling knowledge gaps in other EFDB-relevant variables will require
more effort, but this effort is warranted given their importance for
estimating forest C stock chnages. Although aboveground biomass is the
most studied variable considered here (Table 2) and is the target of
satellite missions
\citep{dubayah_global_2020, quegan_european_2019, nisar_nasaisro_2018},
significant ground-based research effort is required to create accurate
global maps of forest biomass and changes therein
\citep{duncanson_importance_2019, labrière_forest_2023}. Given
observations of increasing tree mortality in some forested regions
\citep{mcdowell_pervasive_2020}, better characterization of forest dead
wood will be critical. Additionally, C stocks in forest organic horizons
and soils can be quite substantial and highly uncertain in many parts of
the world \citep{tifafi_large_2018}. Significant investment in
ground-based forest research will be critical to filling these gaps.

@@

We recommend that, in order to make research valuable to estimate C
stock changes according to methods provided in the IPCC guidelines,
researchers calculate and report results according to IPCC good practice
(Table 3). It is particularly noteworthy that simple decisions on the
presentation of results will determine whether the records meet the
criteria for inclusion in EFDB. Some examples are as follows: (1)
presenting data only in a figure makes them ineligible for inclusion in
EFDB, whereas presentation in a table or supplementary data file allows
inclusion; (2) direct presentation of all relevant variables allows
inclusion, whereas presenting only components of variables of interest
(e.g., parsing litter into fine woody debris, OL, OF, and OH layers) or
requiring simple mathematical operations to obtain a variable of
interest (e.g., \emph{delta.agb} = \emph{ANPP\_woody} -
\emph{woody.mortality.agb}) disqualifies records from inclusion; (3)
matching IPCC-defined thresholds for defining C pools (Table 1), which
may vary by country, can make the data far more relevant for IPCC
accounting (e.g., using a 10 cm cutoff between dead wood and litter,
presenting soil C to a depth of 30 cm). It should also be emphasized
that reporting of 95\% confidence intervals (or other metrics of error),
when applicable, is highly desirable and makes the data more relevant to
IPCC.
